---
title: Tidyeval
author: "Ian Lyttle, adapted from material by Lionel Henry and Hadley Wickham"
output: 
  learnr::tutorial:
    df_print: tibble
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library("learnr")
library("dplyr")
library("tibble")
library("glue")
library("learnr.tidyeval")

options(width = 100)
knitr::opts_chunk$set(comment = "#>")
```

(Note: I think this tutorial is almost ready. I have run out of gas on a few sections and I way want to add a little more supporting material. Otherwise, ready to start showing this around.)

## Introduction

Confession: for years, I have struggled to wrap my head around non-standard evaluation. I would sit down over a weekend and get myself to the point that I _thought_ I understood it, I _seemed_ to be able to write code to use it effectively, then a month later I would have no idea of what I had written or why it seemed to work. Here's a receipt:

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Personal <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> resolution: 2016 will be the year I wrap my head around non-standard evaluation (just like 2015, 2014, 2013, ...)</p>&mdash; Ian Lyttle (@ijlyttle) <a href="https://twitter.com/ijlyttle/status/682678320039899141?ref_src=twsrc%5Etfw">December 31, 2015</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

At first glance, tidyeval can seem just as mysterious - a new way to deal with non-standard evaluation. I came to appreciate that tidyeval is built on a series of principles; as I came to understand (at least some of) these principles, it became easier for me to retrace my steps and to take new ones. At this point I think I understand the basic principles such that, at the risk of telling you more than I know, I can discuss those basics here.

Of course, none of this material is possible without the work done by Lionel Henry and Hadley Wickham, who developed the tidyeval framework. Indeed, the examples shown here are adapted from their sources:

- [Programming with dplyr](http://dplyr.tidyverse.org/articles/programming.html)
- [Tidy evaluation](http://rlang.tidyverse.org/articles/tidy-evaluation.html)
- [Erratum tidyr 0.7.0](https://www.tidyverse.org/articles/2017/09/erratum-tidyr-0.7.0/)

I am particularly impressed with the "Erratum" article, as it explains very well the factors that went into a design choice, as well as the factors that went into the decision to modify that choice.

The goal of this tutorial is to make it easier for you to tackle these references.

As you go through this tutorial, please keep in mind that I am an imperfect interpeter of tidyeval principles. I will do my best to convey what I understand; any shortcomings are my sole responsibility, which I will recitify as they become apparent ([issue page](https://github.com/ijlyttle/learnr.tidyeval/issues)). 

Another thing is that I am purposefully skipping over the theoretical aspects of tidyeval in this tutorial. I have ther greatest respect for the work made by Hadley and Lionel to incorporate quasiquotation theory into tidyeval. It is a mark of that respect and of my lack of familiarity with the theory that I avoid addressing it, and instead enthusiastially refer you to their material.

One more thing (still here?), for the purpose of this tutorial, I have shied away from using the `!!` and `!!!` operators. I have nothing against them; I rather like using them. However, there are a few cases where you can run into trouble, so I decided to leave learning their use as a further exercise to the reader.

One final note - this tutorial is produced using the [__learnr__ package](https://rstudio.github.io/learnr/), which lets you produce a document that includes interactve exercises, and can be published to RStudio's [shinyapps service](http://www.shinyapps.io/).

## What motivates tidyeval?

### Advantages

You are probably used to **dplyr** code that looks like this:

```{r}
filter(starwars, species == "Droid")
```

This does the same thing as base R code: 

```{r}
starwars[starwars$species == "Droid" &  !is.na(starwars$species), ]
```

The dplyr code is more concise - you don't have to repeat the name of the data frame, or specify that you don't want `NA` values.

As well, you may have used dplyr to access databases. If the `starwars` tibble were stored on a remote database, rather than in a local data-frame, dplyr would translate the expression `species == "Droid"` into SQL to obtain the same result. Tidyeval is the framework that dplyr uses to adapt the expression to the situation, so that "it just works".

### Things to keep in mind

However, as [Jar Jar discovered](https://youtu.be/qL1fBDQXSy8?t=1m40s), there is no such thing as a free lunch. There are a couple things to be mindful of.

#### Referential transparency

Most arguments to dplyr functions are not [__referentially transparent__](https://en.wikipedia.org/wiki/Referential_transparency). You can't replace an argument with a seemingly-equivalent argument that you have defined elsewhere.

##### Your turn

Let's look again at our earlier example:

```{r}
filter(starwars, species == "Droid")
```

Try the following substitutions and see what happens:

```{r eval=FALSE}
my_var <- species
filter(starwars, my_var == "Droid")

my_var <- "species"
filter(starwars, my_var == "Droid")
```

```{r starwars-my_var, exercise=TRUE}
filter(starwars, species == "Droid")
```

As you see, neither of these substitutions works how we expect - so it is not obvious how to call dplyr functions without hard-coding everything.

#### Ambiguity

As well, dplyr evaluates expressions ambiguously. As we will see, the code `filter(df, x == y)` can be interpreted differently depending on the data and the context:

```{r eval=FALSE}
df[df$x == df$y, ]
df[df$x == y, ]
df[x == df$y, ]
df[x == y, ]
```

If the "wrong" thing happens when you are working interactively, you can see that there is a problem and correct it. If the "wrong" thing happens when you are calling a function (from a function, from a packge, ...), it can be nearly impossible to figure out what went wrong, let alone fix it.

##### Your turn

Let's say we want to look at the Star Wars characters, comparing eye-color and hair-color.

In the code-block below:

1. Run the code as-is.
1. Uncomment the `eye_color == "black"` line, rerun.
1. Uncomment the `my_color == "black"` line, change the filter expression to `hair_color == my_color`, rerun.
1. Fiddle around with other varaibles and filter expressions.

What can you tell about the rules that `dplyr::filter()` uses to evaluate an expression?

```{r starwars-eye-hair, exercise=TRUE}
# eye_color <- "black"
# my_color <- "black"   
filter(starwars, hair_color == eye_color)
```

### Summary

Tidyeval provides a set of rules for the default evaluation of expressions, and a set of tools that let you change that default whenever you need to. For people who write functions and people who use those functions, tidyeval provides a framework that allows both sets of people be confident that a function will do what they expect it to do.

In the next few sections, we will discuss the principles and concepts of tidyeval, then show how you can use these to call functions that use tidyeval (like those in dplyr). Finally, we will discuss how you can write functions that use the tidyeval framework.

## Principles and concepts

We can demonstrate tidyeval's central idea using strings. Let's say that we want to compose a string to include a name that we specify elsewhere. This warm-up example is lifted from the [dplyr programming vignette](http://dplyr.tidyverse.org/articles/programming.html).

##### Your turn

In the excercise below, you will see that our first attempt is clearly doomed. Comment and un-comment the following lines to see some approaches we can use to compose the string that we expect. Maybe try some compoistions of your own...

In your own words, why does the first attempt fail, while the other two succeed?

```{r warm-up, exercise=TRUE}
name <- "not Hadley"

"How do you do, name?"

# paste0("How do you do, ", name, "?")

# library("glue")
# glue("How do you do, {name}?")
```

In the first attempt, we cannot evaluate the name because it is "quoted": the string itself is simply a value. In the second attempt, we build up the string piece-by-piece: `paste0()` evaluates `name` within the larger set of expressions. In the third attempt we use the __glue__ package, which allows us to compose a string by "unquoting": `{name}` is replaced with value of `name`, using the context from which `glue()` is called.

It may be useful to establish some vocabulary terms: symbols, expressions, and quosures.

### Symbols, expressions, and quosures (oh my!)

This may be a good time to recall the [famous quote from _Advanced R_](https://adv-r.hadley.nz/functions.html#all-calls):

> “To understand computations in R, two slogans are helpful:
> 
> Everything that exists is an object.
> Everything that happens is a function call."
> — John Chambers

The ultimate result of evaluation (tidy or otherwise) is a value. Here, we will go over the objects we use in tidyeval to get to values. Types of values are logicals, doubles, integers, characters, and so on. With a value, everything has been evaluated - there is nothing left to define. Here's an example:

```{r}
is_droid <- starwars$species == "Droid" & !is.na(starwars$species)

is_droid
```

You can send these values as an argument to dplyr's `filter()` function:

```{r}
filter(starwars, is_droid)
```

As you can see, `filter()` is perfectly happy to take this set of values and return the data frame - but that's not why dplyr functions are so useful. We are accustomed to using dplyr like this:

```{r}
filter(starwars, species == "Droid")
```

Here, we used an __expression__, `species == "Droid"`; it is then `filter()`'s job to evaluate this expression into values - then use those values to return the corresponding rows of the data frame. It may be useful to think of an expression as a set of instructions.

As we have found out, when dplyr evaluates an expression, its default is to look for variable-names first in the data frame, then in the context from which you called the function. Going forward, we will go through some examples of how you can use tidyeval to build expressions that are evaluated according to _your_ specifications.

You can compose expressions using __symbols__; in this case, `species` is a symbol. Let's say you are building a shiny app where you need to call a dplyr function on a data frame that you upload to the app. When you are writing the app, you cannot possibly know the variable names that will be used at runtime. 

It may not be evident how to build expression for dplyr to evaluate. We will see how the tidyeval framework lets you build expressions using symbols that you can create at runtime.

If you are already a little familiar with tidyeval, you have heard about __quosures__. These are the central feature of tidyeval. However, I'm not going to say too much about quosures just yet. In the Star Wars movies, Yoda was not introduced until the second movie was released, even though he is a pivotal character.

For now, it may be useful to think about a quosure as an expression that carries along with it the context where it was created. If we think of an expression as a set of instructions, we can think of a quosure as a set of instructions bundled with a set of materials. We will look at quosures more explicitly later in this tutorial.

> "Patience, for the Jedi it is time to eat as well." - a certain Jedi Master

## Tidyeval and tidyverse 

One of the mantras of the tidyverse is "data first!" 

This commonly refers to design of tidyverse functions where a data-frame is the first argument; this makes it easy to use the pipe operator. For example, let's find the youngest character(s) from each species.

```{r}
birth_year <- c(15, 19) # what if filter() did not look at data first?

starwars %>%
  select(name, species, birth_year) %>%
  group_by(species) %>%
  filter(birth_year == min(birth_year, na.rm = TRUE)) 
```

As you can see, we can interpret "data first!" another way: when a tidyverse function evaluates an expression, it looks in the _data first_, then the calling context. We can work confidently with tidyverse functions because they behave consistently according to a convention.

In the remainder of this tutorial, we are going to use tidyeval with dplyr functions: first calling them directly, then writing functions that call dplyr functions. One thing to keep in mind going forward, we are restricting ourselves to using these functions when the data-source is a local data frame. I understand that these priciples extend to using other data sources such as remote databases, but I am not your best guide to these other situations.

## Using functions that use tidyeval

One use for tidyeval is to help you write package functions. Another application is writing __shiny__ apps. 

In a shiny app, let's say you want your user to be able to upload a data-frame, filter on some condition, then make a scatterplot, letting your user specify your axes. Let's also say that you want to use the tidyverse. 

For the filtering step, one challenge may be to compose a filtering condition without knowing what the data-frames's variables are. 

For the visualisation step, __ggplot2__ does not currently use tidyeval; instead you use the `aes_string()` function. But, by becoming familiar with tidyeval now, it will be that much easier to use the tidyeval framework in ggplot2 when it arrives.

Let's pick up the filtering example. As you know, the `dplyr::filter()` function takes a series of expressions that it evaluates to `TRUE` or `FALSE` to determine which rows to return. Let's say that you have a shiny app where you have:

- an input that specifies the variable for the filter - this will be a character string
- an input that specifies the value - this will be a numeric

For the sake of simplicity, let's say that it is single condition and that we are testing equality. For the sake of even more simplicity, let's start out by specifying we will use the `starwars` data; we will filter on `birth_year`, we will vary only the value. 

```{r}
value <- 19

filter(starwars, birth_year == value)
```

So far so good - `filter()` looks for `birth_year` in the data and finds it, then looks for `value` in the data, does not find it, then finds `value` in the context.

You will need to use a little imagination because this is a contrived example. Let's imagine that we don't know that `starwars` does not have a column named `value`. We want to make this evaluation "airtight", we want to make sure that we use `value` from our context.

Let's look back at our string example:

```{r}
library("glue")

value <- 19

glue("We want rows where the birth_year is {value}.")
```

Recall that we are composing a string by unquoting: `{name}` is replaced with value of `name`, using the context from which `glue()` is called. 

Instead of composing strings we need to compose expressions, so Tidyeval provides an __unquoting__ function.
The `UQ()` function acts on expressions in tidyeval like the `{}` acts on strings in `glue()`: it evaluates its contents and incorporates the result.

##### Your turn

Make sure this works:

```{r filter-uq-value, exercise=TRUE}
value <- 19

filter(starwars, birth_year == UQ(value))
```

To convince yourself that the code above is resistant to mischeif, change both instances of `value` to `birth_date` and rerun.

There is another thing we can do here, instead of evaluating the filter statement then printing out the data, we can capture the set of instructions, the expression, and print that instead. To do that, we use the `quo()` function.

```{r filter-uq-value-quo, exercise=TRUE}
value <- 19

quo(
  filter(starwars, birth_year == UQ(value))
)
```

Run this code to see what happens. Switch `value` with `birth_year` again, if you like.

We see again, this word "quosure". Again, let's think about a quosure as an expression (set of instructions) with an attached context (set of materials). We see the instructions (`~filter(starwars, birth_year == 19)`), and we are given an indication of the materials (`local`). For now, we can use the `quo()` function to verify that our instructions are what we expect.

-----

Our next step is to specify the dataframe and the column to filter. Recall that symbols can be used to compose expressions. Tidyeval provides a function to convert strings into symbols, `sym()`. 

```{r}
sym("foo")
```

We can use the `sym()` function to help build the left-hand-side of our expression.

##### Your turn

This code does not work... yet.

```{r filter-sym, exercise=TRUE}
df <- starwars
var_name <- "birth_date"
value <- 19

quo(
  filter(starwars, sym(var_name) == UQ(value))
)
```

As you can see, this doesn't look right. If only there were a way that we could evaluate `sym(var_name)` and incorporate the result into the expression...

Also notice that we don't have to do anything special with the first argument to `filter()`. We can replace the argument with an equivalent object and get the same result; `filter()` is __referentially transparent__ for this argument.

Finally, if you like, remove the `quo()` to make sure it still returns the data we expect.

-----

##### Your turn

There is one more thing we can do to make things airtight. Consider this bit of mischeif:

```{r filter-mischeif, exercise=TRUE}
df <- starwars
var_name <- "parsec"
# parsec <- 12
value <- 19

filter(df, UQ(sym(var_name)) == UQ(value))
```

Run this as is, and we get an error-message that `parsec` is not found. Clearly, it is not a column in the data-frame, and it is not in our context. But what if it were in our context? Uncomment the `parsec <- 12` line and rerun. What happens? Why?

-----

Tidyeval has tool to make sure that we look in _only_ the data for a variable, a pronoun named `.data` to stand in for the data argument. You can use this to compose expressions that throw an error if variables are not found in the data.

Here are some examples:

```{r error=TRUE}
filter(starwars, .data$birth_year == 19)

filter(mtcars, .data$birth_year == 19)

filter(starwars, .data[["birth_year"]] == 19)
```

##### Your turn

Modify the left-hand-side of the filtering expression to use the `.data` pronoun with `var_name`. Make sure that we throw an error regardless of the `parsec <- 12` line being uncommented or not.

```{r filter-mischeif-managed, exercise=TRUE}
df <- starwars
var_name <- "parsec"
# parsec <- 12
value <- 19

filter(df, UQ(sym(var_name)) == UQ(value))
```

When we use tidyeval with dplyr functions, dplyr evaluates the expressions that you provide. It can be useful to keep in mind what the dplyr function is evaluating *to*. `filter()` evaluates its expressions to logical values - it wants `TRUE` or `FALSE`. Some dplyr functions do things a little differently, let's look at some of those differences.   

### "selecty" functions

There is a family of functions that use the __tidyselect__ package as their back-end; `dplyr::select()` is the most prominent. Other "selecty" functions include `tidyr::gather()` and `dplyr::pull()`. They all work the same way because they all need to identify a set of columns within an existing data source.

This excellent [tidyr article](https://www.tidyverse.org/articles/2017/09/erratum-tidyr-0.7.0/) dives into the theory and design choices. There are two big points here.

1. Selecty functions evaluate its expressions to integers that indicate column-positions in the data source.

2. It is best-practice to provide expressions that make sense only in the data, not in the surrounding context.

Let's look at some examples of these principles in action. The usual way to call a select function is just to use the bare column-names, which we now know as symbols. 

```{r}
select(starwars, name, species : starships)
```

We get an identical result if we specify the column-positions explicitly.

```{r eval=FALSE}
select(starwars, 1, 10 : 13)
```

If we want to insist that we look only in the data for the name (not a bad idea), we can use the `.data` pronoun:

```{r eval=FALSE}
select(starwars, .data$name, .data$species : .data$starships)
```

Here's where things can get potentially difficult. Let's say that you want to select from a column-position to until the last column. This works:

```{r eval=FALSE}
select(starwars, .data$name, .data$species : ncol(starwars))
```

But the expression `ncol(starwars)` refers to the context where `starwars` is found. Our best-practice is to refer only to the data. We now have two references to `starwars` in the function call; we would prefer to have only one.

The __tidyselect__ package gives us a solution, a helper function called `last_col()` (if functions had hex-stickers, I imagine this hex-sticker would feature a pint glass, but I digress).

```{r}
library("tidyselect")
select(starwars, .data$name, .data$species : last_col())
```

##### Your turn

Try out some of the techniques featured above to convince yourself that they work.

```{r selecty, exercise=TRUE}
library("tidyselect")
select(starwars, name, species : starships)
```

Try some crazy ideas and see what works. For your craziest idea that works, explain why it works and why it might not be a good idea to use it in practice.

### "namey" functions

Some dplyr functions, in addition to evaluating expressions, assign column names. These include `mutate()`, `summarise()`, and `transmute()`.

For an example, let's find the minimum `birth_year` for each species.

```{r}
starwars %>%
  group_by(species) %>%
  summarise(birth_year_min = min(birth_year, na.rm = TRUE)) 
```

Now I want to parameterise this, in case I want to do the same thing with height. (You are likely suspecting that we should really be making a function to do this - which we will in the next section of this tutorial.) 

Here's one way to do it:

```{r}
var_to_min <- "birth_year"

var_min <- paste0(var_to_min, "_min")

starwars %>%
  group_by(species) %>%
  summarise(UQ(sym(var_min)) := min(UQ(sym(var_to_min)), na.rm = TRUE)) 
```

The first step here is to create a new variable-name by appending `"_min"` to the variable-name of interest. The second step is to adapt the `summarise()` expression. Note that it has both a left side (new name) and a right side (value). The good news is that we can use our unquoting tools on both sides. However, the `=` operator does not support unquoting on the left-side, so we use the `:=` operator from the __rlang__ package - designed just for this purpose. 

The use of the `:=` operator is the thing that we need to keep in mind when using "namey" functions. 

One other point, because the left-side is going to end up as a name, you can unquote a symbol, `UQ(sym(var_min))`, or a string, `UQ(var_min)`.

##### Your turn

Evaluate this solution for yourself:

```{r namey-summarise, exercise=TRUE}
var_to_min <- "birth_year"

var_min <- paste0(var_to_min, "_min")

starwars %>%
  group_by(species) %>%
  summarise(UQ(sym(var_min)) := min(UQ(sym(var_to_min)), na.rm = TRUE)) 
```

Experiment with unquoting the string rather than the symbol, try using different variable names, wrap the call in `quo()` to see the instructions. Have fun.

## Writing functions that use tidyeval

I imagine that there may have been a few times in the previous section when you thought to yourself, "why don't you write a function?!?" 

Yes - we should be writing functions; tidyeval's superpower is that it allows us to write functions that we can be confident will work _how we intend_, after we release them into the world. For this tutorial, I thought to try to introduce one tidyeval component at a time, so I  saved functions for later. You, dear reader, may judge me on that.

Now that we're here, let's look at writing functions.

Recall our example where we filtered the `starwars` dataset based on a string and a numeric value.

```{r filter-str-ex, eval=FALSE}
df <- starwars
var_name <- "birth_date"
value <- 19

filter(starwars, UQ(sym(var_name)) == UQ(value))
```

The rest of these examples are in "interactive" mode so that you can experiment with them. Our first step is to wrap this code in a function.

```{r filter-str-function, exercise=TRUE}
filter_name_value <- function(.data, var_name, value) {
  dplyr::filter(.data, UQ(sym(var_name)) == UQ(value))  
}

filter_name_value(starwars, "birth_year", 19)
```

Try using this function to filter other data frames, for example, `filter_name_value(mtcars, "cyl", 6)`.

### Package considerations

As you know, coding conventions for a package function are a little different than for interactive programming. Most prominently, calls to functions in other packages specify the namespace, i.e. `dplyr::filter()` rather than `filter()`. For commonly-used low level functions, like `UQ()` and `sym()`, from __rlang__, we import these into the package itself. For example, see [`utils-tidy-eval.R`](https://github.com/ijlyttle/learnr.tidyeval/blob/master/R/utils-tidy-eval.R) in the package's R directory.  

As of the writing of this tutorial, the tidyverse folks are working on a new package [__usethis__](https://github.com/r-lib/usethis), that will make it very easy for you to use tidyeval functions in your package-functions, _and_ pass R CMD CHECK. The `use_tidyeval()` function, used to generate `utils-tidy-eval.R`, is not on CRAN yet (2017-11-01), but I imagine it will be there soon.

To demonstrate that you can use these steps to write a function that works and passes R CMD CHECK, I have included a function `ex_filter_name_value()` in this package, you can inspect it [here](). Try it out, i.e. `ex_filter_name_value(mtcars, "cyl", 6)`.

### Quosures

We have written a function that seems to work for a variety of cases, let's work to make this code look-and-feel a little more like "dplyr". We would like to use a bare symbol, rather than a string.

Consider the function `filter_quo_value()` in the exercise below. The operative expression is

```r
UQ(var_quo) == UQ(value)
```

On the right-hand side, inside the `UQ()`, we want something that evaluates to a value. On the left-hand side, inside the `UQ()`, we want something that evaluates to an expression (a set of instructions). 

We have a tool for encapsulating an expression: `quo()`, which builds a quosure. In this case, we are not using (as far as I know) the context (set of materials) part of the quosure, only the expression. 

To use our function, we wrap the bare name of our column in `quo()`, in this example: `quo(birth_year)`.

Like the example above, experiment with this function.

```{r filter-quo-function, exercise=TRUE}
filter_quo_value <- function(.data, var_quo, value) {
  dplyr::filter(.data, UQ(var_quo) == UQ(value))  
}

filter_quo_value(starwars, quo(birth_year), 19)
```

We are closer, but we still have this `quo()` hanging around; it's not so convenient to remember to use when we call the function. It would be great if we our function could just handle the quoting for us. This is the job of the `enquo()` function.

We use `enquo()` in our function to capture the argument as a quosure. Try it out:

```{r filter-sym-value, exercise=TRUE}
filter_sym_value <- function(.data, sym, value) {
  
  sym <- enquo(sym)
  
  dplyr::filter(.data, UQ(sym) == UQ(value)) 
}

filter_sym_value(starwars, birth_year, 19)
```

Bonus task: amend the exercise above to make a call to `filter_sym_value()` using a string, `"birth_year"`. Do not change the function, change the call. Look earlier in this tutorial for a hint.

### Renaming variables

Recall this exercise:

```{r eval=FALSE}
var_to_min <- "birth_year"

var_min <- paste0(var_to_min, "_min")

starwars %>%
  group_by(species) %>%
  summarise(UQ(sym(var_min)) := min(UQ(sym(var_to_min)), na.rm = TRUE)) 
```

Let's rewrite this as a function where we want to specify the variable that we minimise and the variable that use for grouping.

Talk about `quo_name()`, remind about `:=`.

```{r renaming-variables, exercise=TRUE}
summarise_min <- function(.data, var_min, var_group) {
  
  var_group <- enquo(var_group)
  var_min <- enquo(var_min)
  
  name_var_min <- rlang::quo_name(var_min)
  name_var_new <- paste0(name_var_min, "_min")
  
  .data %>%
    dplyr::group_by(UQ(var_group)) %>%
    dplyr::summarise(UQ(name_var_new) := min(UQ(var_min), na.rm = TRUE))
}

summarise_min(starwars, birth_year, species)
```

### Multiple variables

Let's say we want to specify more than one grouping variable. Talk about using `quos()` to capture the `...` into a list of quosures. Talk about using `UQS()` to unquote a list of quosures.

```{r multiple-variables, exercise=TRUE}
summarise_more_min <- function(.data, var_min, ...) {
  
  var_group <- quos(...)
  var_min <- enquo(var_min)
  
  name_var_min <- rlang::quo_name(var_min)
  name_var_new <- paste0(name_var_min, "_min")
  
  .data %>%
    dplyr::group_by(UQS(var_group)) %>%
    dplyr::summarise(UQ(name_var_new) := min(UQ(var_min), na.rm = TRUE))
}

summarise_more_min(starwars, birth_year, species, hair_color)
```

## More stuff

Hopefully, you have been able to use this tutorial to get a sense of tidyeval principles and tools. I would encourage you to go through the three main references by the tidyeval architects, you can use these to deepen your knowledge ,and introduce yourself to the supporting theory:

- [Programming with dplyr](http://dplyr.tidyverse.org/articles/programming.html)
- [Tidy evaluation](http://rlang.tidyverse.org/articles/tidy-evaluation.html)
- [Erratum tidyr 0.7.0](https://www.tidyverse.org/articles/2017/09/erratum-tidyr-0.7.0/)

There are also a lot of resources appearing on around the internet, produced by other R users who are also wrapping their heads around tidyeval. Mara Averick has the canonical [tidyeval resource roundup](https://maraaverick.rbind.io/2017/08/tidyeval-resource-roundup/), so the best thing I can do here is to point you that-a-way.



