---
title: Tidyeval
author: "Ian Lyttle, adapted from material by Lionel Henry and Hadley Wickham"
output: 
  learnr::tutorial:
    df_print: tibble
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library("learnr")
library("dplyr")
library("tibble")
library("glue")
library("learnr.tidyeval")

options(width = 100)
knitr::opts_chunk$set(comment = "#>")
```

## Introduction

Confession: for years, I have struggled to wrap my head around non-standard evaluation. I would sit down over a weekend and get myself to the point that I _thought_ I understood it, I _seemed_ to be able to write code to use it effectively, then a month later I would have no idea of what I had written or why it seemed to work. Here's a receipt:

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Personal <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> resolution: 2016 will be the year I wrap my head around non-standard evaluation (just like 2015, 2014, 2013, ...)</p>&mdash; Ian Lyttle (@ijlyttle) <a href="https://twitter.com/ijlyttle/status/682678320039899141?ref_src=twsrc%5Etfw">December 31, 2015</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

At first glance, tidyeval can seem just as mysterious - a new way to deal with non-standard evaluation. I came to appreciate that tidyeval is built on a series of principles; as I came to understand (at least some of) these principles, it became easier for me to retrace my steps and to take new ones. At this point I think I understand the basic principles such that, at the risk of telling you more than I know, I can discuss those basics here.

Of course, none of this material is possible without the work done by Lionel Henry and Hadley Wickham, who developed the tidyeval framework. Indeed, the examples shown here are adapted from their sources:

- [Programming with dplyr](http://dplyr.tidyverse.org/articles/programming.html)
- [Tidy evaluation](http://rlang.tidyverse.org/articles/tidy-evaluation.html)
- [Erratum tidyr 0.7.0](https://www.tidyverse.org/articles/2017/09/erratum-tidyr-0.7.0/)

I am particularly impressed with the "Erratum" article, as it explains very well the factors that went into a design choice, as well as the factors that went into the decision to modify that choice.

The goal of this tutorial is to make it easier for you to tackle these references.

As you go through this tutorial, please keep in mind that I am an imperfect interpeter of tidyeval principles. I will do my best to convey what I understand; any shortcomings are my sole responsibility, which I will recitify as they become apparent ([issue page](https://github.com/ijlyttle/learnr.tidyeval/issues)).

## What motivates tidyeval?

### Advantages

You are probably used to **dplyr** code that looks like this:

```{r}
filter(starwars, species == "Droid")
```

This does the same thing as base R code: 

```{r}
starwars[starwars$species == "Droid" &  !is.na(starwars$species), ]
```

The dplyr code is more concise - you don't have to repeat the name of the data frame, or specify that you don't want `NA` values.

As well, you may have used dplyr to access databases. If the `starwars` tibble were stored on a remote database, rather than in a local data-frame, dplyr would translate the expression `species == "Droid"` into SQL to obtain the same result. Tidyeval is the framework that dplyr uses to adapt the expression to the situation, so that "it just works".

### Things to keep in mind

However, as [Jar Jar discovered](https://youtu.be/qL1fBDQXSy8?t=1m40s), there is no such thing as a free lunch. There are a couple things to be mindful of.

#### Referential transparency

Most arguments to dplyr functions are not [__referentially transparent__](https://en.wikipedia.org/wiki/Referential_transparency). You can't replace an argument with a seemingly-equivalent argument that you have defined elsewhere.

##### Your turn

Let's look again at our earlier example:

```{r}
filter(starwars, species == "Droid")
```

Try the following substitutions and see what happens:

```{r eval=FALSE}
my_var <- species
filter(starwars, my_var == "Droid")

my_var <- "species"
filter(starwars, my_var == "Droid")
```

```{r starwars-my_var, exercise=TRUE}
filter(starwars, species == "Droid")
```

As you see, neither of these substitutions works how we expect - so it is not obvious how to call dplyr functions without hard-coding everything.

#### Ambiguity

As well, dplyr evaluates expressions ambiguously. As we will see, the code `filter(df, x == y)` can be interpreted differently depending on the data and the context:

```{r eval=FALSE}
df[df$x == df$y, ]
df[df$x == y, ]
df[x == df$y, ]
df[x == y, ]
```

If the "wrong" thing happens when you are working interactively, you can see that there is a problem and correct it. If the "wrong" thing happens when you are calling a function (from a function, from a packge, ...), it can be nearly impossible to figure out what went wrong, let alone fix it.

##### Your turn

Let's say we want to look at the Star Wars characters, comparing eye-color and hair-color.

In the code-block below:

1. Run the code as-is.
1. Uncomment the `eye_color == "black"` line, rerun.
1. Uncomment the `my_color == "black"` line, change the filter expression to `hair_color == my_color`, rerun.
1. Fiddle around with other varaibles and filter expressions.

What can you tell about the rules that `dplyr::filter()` uses to evaluate an expression?

```{r starwars-eye-hair, exercise=TRUE}
# eye_color <- "black"
# my_color <- "black"   
filter(starwars, hair_color == eye_color)
```

### Summary

Tidyeval provides a set of rules for the default evaluation of expressions, and a set of tools that let you change that default whenever you need to. For people who write functions and people who use those functions, tidyeval provides a framework that allows both sets of people be confident that a function will do what they expect it to do.

In the next few sections, we will discuss the principles and concepts of tidyeval, then show how you can use these to call functions that use tidyeval (like those in dplyr). Finally, we will discuss how you can write functions that use the tidyeval framework.

## Principles and concepts

We can demonstrate tidyeval's central idea using strings. Let's say that we want to compose a string to include a name that we specify elsewhere. This warm-up example is lifted from the [dplyr programming vignette](http://dplyr.tidyverse.org/articles/programming.html).

##### Your turn

In the excercise below, you will see that our first attempt is clearly doomed. Comment and un-comment the following lines to see some approaches we can use to compose the string that we expect. Maybe try some compoistions of your own...

In your own words, why does the first attempt fail, while the other two succeed?

```{r warm-up, exercise=TRUE}
name <- "not Hadley"

"How do you do, name?"

# paste0("How do you do, ", name, "?")

# library("glue")
# glue("How do you do, {name}?")
```

In the first attempt, we cannot evaluate the name because it is "quoted": the string itself is simply a value. In the second attempt, we build up the string piece-by-piece: `paste0()` evaluates `name` within the larger set of expressions. In the third attempt we use the __glue__ package, which allows us to compose a string by "unquoting": `{name}` is replaced with value of `name`, using the context from which `glue()` is called.

It may be useful to establish some vocabulary terms: symbols, expressions, and quosures.

### Symbols, expressions, and quosures (oh my!)

This may be a good time to recall the [famous quote from _Advanced R_](https://adv-r.hadley.nz/functions.html#all-calls):

> “To understand computations in R, two slogans are helpful:
> 
> Everything that exists is an object.
> Everything that happens is a function call."
> — John Chambers

The ultimate result of evaluation (tidy or otherwise) is a value. Here, we will go over the objects we use in tidyeval to get to values. Types of values are logicals, doubles, integers, characters, and so on. With a value, everything has been evaluated - there is nothing left to define. Here's an example:

```{r}
is_droid <- starwars$species == "Droid" & !is.na(starwars$species)

is_droid
```

You can send these values as an argument to dplyr's `filter()` function:

```{r}
filter(starwars, is_droid)
```

As you can see, `filter()` is perfectly happy to take this set of values and return the data frame - but that's not why dplyr functions are so useful. We are accustomed to using dplyr like this:

```{r}
filter(starwars, species == "Droid")
```

Here, we used an __expression__, `species == "Droid"`; it is then `filter()`'s job to evaluate this expression into values - then use those values to return the corresponding rows of the data frame. It may be useful to think of an expression as a set of instructions.

As we have found out, when dplyr evaluates an expression, its default is to look for variable-names first in the data frame, then in the context from which you called the function. Going forward, we will go through some examples of how you can use tidyeval to build expressions that are evaluated according to _your_ specifications.

You can compose expressions using __symbols__; in this case, `species` is a symbol. Let's say you are building a shiny app where you need to call a dplyr function on a data frame that you upload to the app. When you are writing the app, you cannot possibly know the variable names that will be used at runtime. 

It may not be evident how to build expression for dplyr to evaluate. We will see how the tidyeval framework lets you build expressions using symbols that you can create at runtime.

If you are already a little familiar with tidyeval, you have heard about __quosures__. These are the central feature of tidyeval. However, I'm not going to say too much about quosures just yet. In the Star Wars movies, Yoda was not introduced until the second movie was released, even though he is a pivotal character.

For now, it may be useful to think about a quosure as an expression that carries along with it the context where it was created. If we think of an expression as a set of instructions, we can think of a quosure as a set of instructions bundled with a set of materials. We will look at quosures more explicitly later in this tutorial.

> "Patience, for the Jedi it is time to eat as well." - a certain Jedi Master

## Tidyeval and tidyverse 

One of the mantras of the tidyverse is "data first!" 

This commonly refers to design of tidyverse functions where a data-frame is the first argument; this makes it easy to use the pipe operator. For example, let's find the youngest character(s) from each species.

```{r}
birth_year <- c(15, 19) # what if filter() did not look at data first?

starwars %>%
  select(name, species, birth_year) %>%
  group_by(species) %>%
  filter(birth_year == min(birth_year, na.rm = TRUE)) 
```

As you can see, we can interpret "data first!" another way: when a tidyverse function evaluates an expression, it looks in the _data first_, then the calling context. We can work confidently with tidyverse functions because they behave consistently according to a convention.

In the remainder of this tutorial, we are going to use tidyeval with dplyr functions: first calling them directly, then writing functions that call dplyr functions. One thing to keep in mind going forward, we are restricting ourselves to using these functions when the data-source is a local data frame. I understand that these priciples extend to using other data sources such as remote databases, but I am not your best guide to these other situations.

## Using functions that use tidyeval

One use for tidyeval is to help you write package functions. Another application is writing __shiny__ apps. 

In a shiny app, let's say you want your user to be able to upload a data-frame, filter on some condition, then make a scatterplot, letting your user specify your axes. Let's also say that you want to use the tidyverse. 

For the filtering step, one challenge may be to compose a filtering condition without knowing what the data-frames's variables are. 

For the visualisation step, __ggplot2__ does not currently use tidyeval; instead you use the `aes_string()` function. But, by becoming familiar with tidyeval now, it will be that much easier to use the tidyeval framework in ggplot2 when it arrives.

Let's pick up the filtering example. As you know, the `dplyr::filter()` function takes a series of expressions that it evaluates to `TRUE` or `FALSE` to determine which rows to return. Let's say that you have a shiny app where you have:

- an input that specifies the variable for the filter - this will be a character string
- an input that specifies the value - this will be a numeric

For the sake of simplicity, let's say that it is single condition and that we are testing equality. For the sake of even more simplicity, let's start out by specifying we will use the `starwars` data; we will filter on `birth_year`, we will vary only the value. 

```{r}
value <- 19

filter(starwars, birth_year == value)
```

So far so good - `filter()` looks for `birth_year` in the data and finds it, then looks for `value` in the data, does not find it, then finds `value` in the context.

You will need to use a little imagination because this is a contrived example. Let's imagine that we don't know that `starwars` does not have a column named `value`. We want to make this evaluation "airtight", we want to make sure that we use `value` from our context.

Let's look back at our string example:

```{r}
library("glue")

value <- 19

glue("We want rows where the birth_year is {value}.")
```

Recall that we are composing a string by unquoting: `{name}` is replaced with value of `name`, using the context from which `glue()` is called. 

Instead of composing strings we need to compose expressions, so Tidyeval provides an __unquoting__ function.
The `UQ()` function acts on expressions in tidyeval like the `{}` acts on strings in `glue()`: it evaluates its contents and incorporates the result.

##### Your turn

Make sure this works:

```{r filter-uq-value, exercise=TRUE}
value <- 19

filter(starwars, birth_year == UQ(value))
```

To convince yourself that the code above is resistant to mischeif, change both instances of `value` to `birth_date` and rerun.

There is another thing we can do here, instead of evaluating the filter statement then printing out the data, we can capture the set of instructions, the expression, and print that instead. To do that, we use the `quo()` function.

```{r filter-uq-value-quo, exercise=TRUE}
value <- 19

quo(
  filter(starwars, birth_year == UQ(value))
)
```

Run this code to see what happens. Switch `value` with `birth_year` again, if you like.

We see again, this word "quosure". Again, let's think about a quosure as an expression (set of instructions) with an attached context (set of materials). We see the instructions (`~filter(starwars, birth_year == 19)`), and we are given an indication of the materials (`local`). For now, we can use the `quo()` function to verify that our instructions are what we expect.

-----

Our next step is to specify the dataframe and the column to filter. Recall that symbols can be used to compose expressions. Tidyeval provides a function to convert strings into symbols, `sym()`. 

```{r}
sym("foo")
```

We can use the `sym()` function to help build the left-hand-side of our expression.

##### Your turn

This code does not work... yet.

```{r filter-sym, exercise=TRUE}
df <- starwars
var_name <- "birth_date"
value <- 19

quo(
  filter(starwars, sym(var_name) == UQ(value))
)
```

As you can see, this doesn't look right. If only there were a way that we could evaluate `sym(var_name)` and incorporate the result into the expression...

Also notice that we don't have to do anything special with the first argument to `filter()`. We can replace the argument with an equivalent object and get the same result; `filter()` is __referentially transparent__ for this argument.

Finally, if you like, remove the `quo()` to make sure it still returns the data we expect.

-----

##### Your turn

There is one more thing we can do to make things airtight. Consider this bit of mischeif:

```{r filter-mischeif, exercise=TRUE}
df <- starwars
var_name <- "parsec"
# parsec <- 12
value <- 19

filter(df, UQ(sym(var_name)) == UQ(value))
```

Run this as is, and we get an error-message that `parsec` is not found. Clearly, it is not a column in the data-frame, and it is not in our context. But what if it were in our context? Uncomment the `parsec <- 12` line and rerun. What happens? Why?

-----

Tidyeval has tool to make sure that we look in _only_ the data for a variable, a pronoun named `.data` to stand in for the data argument. You can use this to compose expressions that throw an error if variables are not found in the data.

Here are some examples:

```{r error=TRUE}
filter(starwars, .data$birth_year == 19)

filter(mtcars, .data$birth_year == 19)

filter(starwars, .data[["birth_year"]] == 19)
```

##### Your turn

Modify the left-hand-side of the filtering expression to use the `.data` pronoun with `var_name`. Make sure that we throw an error regardless of the `parsec <- 12` line being uncommented or not.

```{r filter-mischeif-managed, exercise=TRUE}
df <- starwars
var_name <- "parsec"
# parsec <- 12
value <- 19

filter(df, UQ(sym(var_name)) == UQ(value))
```

### "selecty" functions

### "namey" functions

## Writing functions that use tidyeval

pronouns 

## More stuff




