---
title: Tidyeval
author: "Ian Lyttle, adapted from material by Lionel Henry and Hadley Wickham"
output: 
  learnr::tutorial:
    df_print: tibble
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library("learnr")
library("dplyr")
library("tibble")
library("glue")
library("learnr.tidyeval")

tbl_mtcars <- as_tibble(mtcars)
```

## Introduction

A few introductory words and a disclaimer.

Confession: for me, this tutorial is an exercise in self-help. I struggled to wrap my head around non-standard evaluation


<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Personal <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> resolution: 2016 will be the year I wrap my head around non-standard evaluation (just like 2015, 2014, 2013, ...)</p>&mdash; Ian Lyttle (@ijlyttle) <a href="https://twitter.com/ijlyttle/status/682678320039899141?ref_src=twsrc%5Etfw">December 31, 2015</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


## What motivates tidyeval?



## Tidyeval concepts

## Using functions that use tidteval

## Writing functions that use tidyeval

## More stuff

This is going to start out as an adaptation of the __dplyr__ [vignette on programming using tidyeval](http://dplyr.tidyverse.org/articles/programming.html).  

You are probably used to **dplyr** code that looks like this:

```{r tbl_mtcars-nse, exercise=TRUE}
tbl_mtcars %>%
  filter(cyl == 4, gear == 3)
```

Which does the same thing as: 

```{r tbl_mtcars-se, exercise=TRUE}
tbl_mtcars[tbl_mtcars$cyl == 4 & tbl_mtcars$gear == 3]
```

- Operations on data frames

```{r does-not-work, exercise=TRUE}
greet <- function(name) {
  "How do you do, name?"
}

greet("your name here")
```

look here

```{r glue2, exercise=TRUE}
greet <- function(name) {
  glue::glue("How do you do, {name}?")
}

greet("your name here")
```

